CF Registry Analysis

### 2025-09-16 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
updates for Monique

```{r}
#| include: false
library(Hmisc)
library(zoo)
library(tidyverse)
library(childsds)
library(gt)
library(gtsummary)
library(forcats)
library(labelled)
library(survival)
library(survminer)
library(knitr)
library(eulerr)
library(broom)
library(emmeans)
library(ggstats)
library(patchwork)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",

@@

annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,

@@

  "White", "Black or African American", "American Indian or Alaska Native",
  "Asian", "Native Hawaiian or Other Pacific ", "Some other race"

@@

demogs$Race <- factor(demogs$Race,

@@

    "White", "American Indian or Alaska Native", "Asian",
    "Black or African American", "More than one race",
    "Native Hawaiian or Other Pacific ", "Some other race"

@@

    "White", "Other race", "Other race",
    "Black or African American", "More than one race",
    "Other race", "Other race"

@@

demogs$Hispanicrace <- factor(demogs$Hispanicrace,

@@

annual$bmi_perc_cdc <- sds(annual$bmi,

@@

  sex = annual$Gender, male = "M", female = "F",
  item = "bmi", ref = cdc.ref, type = "perc"
) * 100
annual$bmi_perc_who <- sds(annual$bmi,

@@

  sex = annual$Gender, male = "M", female = "F",
  item = "bmi", ref = who.ref, type = "perc"
) * 100

@@

    return(cut(bmi_perc, c(-Inf, 10, 50, 85, Inf),

@@

      return(cut(bmi, c(-Inf, 18.5, 23, 25, Inf),

@@

@@

      return(cut(bmi, c(-Inf, 18.5, 22, 25, Inf),

@@

annual$MutClass <- factor(annual$MutClass,

@@

annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,

@@

  breaks = c(0, 1, 2, 3, Inf), right = F

@@

annual$age_group <- cut(annual$Age_YrEnd,

@@

  "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_pulmonarycomplications1", "A_mycocultureresults3"

@@

  "A_corticosteroids1", "A_feeding2",
  "A_Vx770", "A_VX809comb", bugs, "tube_feeding", "liver_disease"

@@

annual[, yn_vars] <- lapply(annual[, yn_vars], factor,

@@

annual$A_IsOnEnzymes <- factor(annual$A_IsOnEnzymes,
  levels = c(1, 0), labels = c(" Yes", " No")

@@

annual$ppFEV1_cat <- cut(annual$A_FEV1_pct_predicted,
  breaks = c(0, 40, 80, 100, Inf), right = F

@@

annual$ppFEV1_cat <- factor(annual$ppFEV1_cat,

@@

  "age_group", "Gender", "bmi_perc", "bmi_cat", "MutClass", "Race",
  "Hispanicrace", "A_IsOnEnzymes", "ppFEV1_cat", "tube_feeding",
  "NumPulmExacerbation", "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_mycocultureresults3", "A_corticosteroids1", "cfrd_status_annual",
  "A_Vx770", "A_VX809comb", "liver_disease"

@@

  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
  Race = "Race", Hispanicrace = "Hispanic",
  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",

@@

  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",

@@

  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
  height_perc = "Height percentile", bmi_cat = "BMI category",

@@

  A_mycocultureresults3 = "Microorganisms in myco culture", eDWID = "ID",
  tstart = "Interval start", tstop = "Interval stop",
  Age_YrEnd = "Age", bmi_perc = "BMI %ile"

@@

```{r}
# Per reviewer, out of X individuals with a diagnosis of CFRD, X number had an
# OGTT (and X had an A1c, but no OGTT) that year or the year prior.
# This is super slow but idc idc
t <- apply(first_cfrd, 1, function(r) {
  id <- as.numeric(r["eDWID"])
  year <- as.numeric(r["first_cfrd"])
  d <- annual[annual$eDWID == id & annual$ReviewYear %in% c(year, year - 1), ]
  if (any(!is.na(d$A_hgba1c))) {
    a1c <- T
  } else {
    a1c <- F
  }
  if (any(!is.na(d["A_twohour_bloodglucose"]) |
    !is.na(d["A_ogttfast_bloodglucose"]))) {
    ogtt <- T
  } else {
    ogtt <- F
  }
  if (a1c & ogtt) {
    return("both")
  } else if (a1c & !ogtt) {
    return("a1c")
  } else if (ogtt & !a1c) {
    return("ogtt")
  } else if (!ogtt & !a1c) {
    return("neither")
  }
})
```

@@

```{r}
data.frame(data_dict$code_meaning[data_dict$var %in%
  sub("A_", "", liver_fields)]) %>%
  gt() %>%
  tab_options(column_labels.hidden = TRUE)
```

@@

  new_cases_10_18 <- sum(inc$n[inc$CFRD == "CFRD+" &
    inc$age_group == "[10,19)"])
  at_risk_10_18 <- sum(inc$n[inc$age_group == "[10,19)"])
  inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)

  new_cases_19_30 <- sum(inc$n[inc$CFRD == "CFRD+" &
    inc$age_group == "[19,31)"])
  at_risk_19_30 <- sum(inc$n[inc$age_group == "[19,31)"])
  inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)

  new_cases_30 <- sum(inc$n[inc$CFRD == "CFRD+" &
    inc$age_group == "[31, Inf)"])
  at_risk_30 <- sum(inc$n[inc$age_group == "[31, Inf)"])
  inc_30 <- new_cases_30 / (at_risk_30 / py)

@@

```{r}
#| message: false
annual %>%
  filter(ReviewYear %in% c(2004, 2011, 2018)) %>%
  select(
    ReviewYear, CFRD, Age_YrEnd, Gender, Race, Hispanicrace, bmi_perc,
    bmi_cat, MutClass, tube_feeding, liver_disease,
    A_FEV1_pct_predicted, A_FVC_pct_predicted, NumPulmExacerbation,
    A_corticosteroids1, A_IsOnEnzymes, A_Vx770, A_VX809comb
  ) %>%
  tbl_strata(
    strata = ReviewYear,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(
          by = CFRD, missing_text = "Missing",
          type = list(A_corticosteroids1 ~ "categorical"),
          statistic = list(c(
            "Age_YrEnd", "bmi_perc", "A_FEV1_pct_predicted",
            "A_FVC_pct_predicted"
          ) ~ "{mean} ({sd})")
        ) %>%
        add_p(test = list(c(
          "Age_YrEnd", "bmi_perc", "A_FEV1_pct_predicted",
          "A_FVC_pct_predicted"
        ) ~ "t.test")) %>%
        add_overall() %>%
        separate_p_footnotes()
  )
```

@@

```{r}
screening <- annual %>%
  group_by(ReviewYear, age_group) %>%
  summarise(
    `Screened HbA1c` = mean(!is.na(A_hgba1c)),
    `Screened OGTT` = mean(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose)),
    `Screened Either` = mean(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose) | !is.na(A_hgba1c)),
    .groups = "drop"
  )
screening_table <- screening %>%
  select(ReviewYear, age_group, `Screened OGTT`) %>%
  mutate(`Screened OGTT` = round(`Screened OGTT` * 100, 2)) %>%
  pivot_wider(names_from = age_group, values_from = `Screened OGTT`)
kable(screening_table, caption = "% Screened by OGTT")
screening <- screening %>%
  pivot_longer(`Screened HbA1c`:`Screened Either`)
screening_plot <- ggplot(screening, aes(
  x = ReviewYear, y = value,
  color = age_group
)) +
  geom_line() +
  xlab("Year") +
  facet_wrap(~name, nrow = 3) +
  theme_bw() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_discrete(name = "Age Group") +
  theme(text = element_text(size = 20), axis.title.y = element_blank())
screening_plot
ggsave("./Dissemination/Figures/screening_rates.png",
  plot = screening_plot, width = 12, height = 9, units = "in"
)
```

@@

```{r results='asis'}
# How many OGTTs are non-CFRD people getting 2016-2018
ogtts <- annual %>%
  filter(ReviewYear %in% 2016:2018, CFRD == "CFRD-") %>%
  group_by(eDWID) %>%
  summarise(
    `Age Group` = last(na.omit(age_group)),
    `OGTTs` = sum(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose))
  )
# Add demographics
ogtts$eDWID <- as.numeric(ogtts$eDWID)
ogtts <- left_join(ogtts, demogs, by = join_by(eDWID))
label(ogtts$eDWID) <- "ID"
# Format OGTT counts
ogtts$OGTTs <- factor(ogtts$OGTTs,
  levels = 0:3, labels = c(
    "No OGTT", "At least 1 OGTT", "At least 1 OGTT", "At least 1 OGTT"
  )
)
# Make table for each variable
age <- ogtts %>%
  select(`Age Group`, OGTTs) %>%
  tbl_summary(by = `Age Group`) %>%
  add_ci() %>%
  add_p() %>%
  print()
gender <- ogtts %>%
  select(Gender, OGTTs) %>%
  tbl_summary(by = Gender) %>%
  add_ci() %>%
  add_p() %>%
  print()
muts <- ogtts %>%
  select(MutClass, OGTTs) %>%
  tbl_summary(by = MutClass) %>%
  add_ci() %>%
  add_p() %>%
  print()
race <- ogtts %>%
  select(Race, OGTTs) %>%
  tbl_summary(by = Race) %>%
  add_ci() %>%
  add_p() %>%
  print()
eth <- ogtts %>%
  select(Hispanicrace, OGTTs) %>%
  tbl_summary(by = Hispanicrace) %>%
  add_ci() %>%
  add_p() %>%
  print()
```

@@

```{r}
p <- prev_year_age %>%
  pivot_longer(`[10,19)`:`[31, Inf)`, names_to = "Age Group") %>%
  rename(Year = ReviewYear)
p$name <- "Prevalence of CFRD"
i <- incidence %>%
  select(
    Year, `Total Incidence Age 10-18`, `Total Incidence Age 19-30`,
    `Total Incidence Age >= 31`
  ) %>%
  pivot_longer(`Total Incidence Age 10-18`:`Total Incidence Age >= 31`,
    names_to = "Age Group"
  )
i$name <- "Incidence of CFRD"
s <- screening %>%
  filter(name == "Screened OGTT") %>%
  select(ReviewYear, age_group, value) %>%
  rename(Year = ReviewYear, `Age Group` = age_group)
s$name <- "Screening Rates for CFRD"
s$value <- s$value * 100
t <- rbind(p, i, s)
t$`Age Group` <- factor(t$`Age Group`,
  levels = c(
    "[10,19)", "[19,31)", "[31, Inf)", "Total Incidence Age >= 31",
    "Total Incidence Age 10-18", "Total Incidence Age 19-30"
  ),
  labels = c(
    "10-18", "19-30", ">= 31", ">= 31",
    "10-18", "19-30"
  )
)
t <- t[t$Year >= 2004, ]
# Figure 1
fig1a <- ggplot(
  t[t$name == "Screening Rates for CFRD", ],
  aes(x = Year, y = value, linetype = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Percentage") +
  ggtitle("Screening Rates for CFRD")
fig1b <- ggplot(
  t[t$name == "Incidence of CFRD", ],
  aes(x = Year, y = value, linetype = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Per 1,000 Person-Years") +
  ggtitle("Incidence of CFRD")
fig1c <- ggplot(
  t[t$name == "Prevalence of CFRD", ],
  aes(x = Year, y = value, linetype = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Per 1,000 Person-Years") +
  ggtitle("Prevalence of CFRD")
fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect") +
  plot_annotation(title = "Figure 1: Trends in Screening, Incidence, and Prevalence of CFRD from 2003-2018, by age group ")
fig1
ggsave("./Dissemination/Figures/fig1.png",
  plot = fig1, width = 12, height = 3, units = "in", dpi = 900
)
# Graphical abstract is the same but with color
fig1a <- ggplot(
  t[t$name == "Screening Rates for CFRD", ],
  aes(x = Year, y = value, color = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Percentage") +
  ggtitle("Screening Rates for CFRD")
fig1b <- ggplot(
  t[t$name == "Incidence of CFRD", ],
  aes(x = Year, y = value, color = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Per 1,000 Person-Years") +
  ggtitle("Incidence of CFRD")
fig1c <- ggplot(
  t[t$name == "Prevalence of CFRD", ],
  aes(x = Year, y = value, color = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Per 1,000 Person-Years") +
  ggtitle("Prevalence of CFRD")
fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect")+
  plot_annotation(title = "US Cystic Fibrosis Foundation Patient Registry Trends in Screening, Incidence, and Prevalence of Cystic Fibrosis Related Diabetes from 2003-2018")
fig1
ggsave("./Dissemination/Figures/graphical_abstract.png",
  plot = fig1, width = 13, height = 3, units = "in", dpi = 900
)
```

@@

```{r results='asis'}
f_vx770 <- update(f, . ~ . + A_Vx770-A_corticosteroids1-NumPulmExacerbation-age_group)
vx770_cox_res <- coxph(f_vx770,
  data = annual[annual$ReviewYear >= 2011, ],
  cluster = eDWID
)
t1 <- vx770_cox_res %>%
  tbl_regression(exponentiate = T)
t1
ggcoef_table(vx770_cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

@@

```{r}
f_vx809comb <- update(f, . ~ . + A_VX809comb-A_corticosteroids1-NumPulmExacerbation-age_group-Hispanicrace)
vx809comb_cox_res <- coxph(f_vx809comb,
  data = annual[annual$ReviewYear >= 2015, ],
  cluster = eDWID
)
t2 <- vx809comb_cox_res %>%
  tbl_regression(exponentiate = T)
t2
ggcoef_table(vx809comb_cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

@@

```{r}
t1 <- era1 %>% tbl_regression(exponentiate = T)
t1
ggcoef_model(era1,
  show_p_values = F, add_reference_rows = T, exponentiate = T, shape_guide = F,
  signif_stars = F, colour = "significance", shape = "reference_row",
  shape_values = c(16, 15)
)
```

>>>>

```{r}
#| include: false
library(Hmisc)
library(zoo)
library(tidyverse)
library(childsds)
library(gt)
library(gtsummary)
library(forcats)
library(labelled)
library(survival)
library(survminer)
library(knitr)
library(eulerr)
library(broom)
library(emmeans)
library(ggstats)
library(patchwork)
home_dir <- switch(
    Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
    "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
github_dir <- switch(
    Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Documents/GitHub",
    "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

annual <- read.csv(
    "./Data_Raw/CFF19_Annualized_Del1.csv",

@@

annual$cfrd_status_annual <- factor(
    annual$cfrd_status_annual,

@@

    "White",
    "Black or African American",
    "American Indian or Alaska Native",
    "Asian",
    "Native Hawaiian or Other Pacific ",
    "Some other race"

@@

demogs$Race <- factor(
    demogs$Race,

@@

        "White",
        "American Indian or Alaska Native",
        "Asian",
        "Black or African American",
        "More than one race",
        "Native Hawaiian or Other Pacific ",
        "Some other race"

@@

        "White",
        "Other race",
        "Other race",
        "Black or African American",
        "More than one race",
        "Other race",
        "Other race"

@@

demogs$Hispanicrace <- factor(
    demogs$Hispanicrace,

@@

annual$bmi_perc_cdc <- sds(
    annual$bmi,

@@

    sex = annual$Gender,
    male = "M",
    female = "F",
    item = "bmi",
    ref = cdc.ref,
    type = "perc"
) *
    100
annual$bmi_perc_who <- sds(
    annual$bmi,

@@

    sex = annual$Gender,
    male = "M",
    female = "F",
    item = "bmi",
    ref = who.ref,
    type = "perc"
) *
    100

@@

        return(cut(
            bmi_perc,
            c(-Inf, 10, 50, 85, Inf),

@@

            return(cut(
                bmi,
                c(-Inf, 18.5, 23, 25, Inf),

@@

                    "Undernourished",
                    "At Risk",
                    "Target/Goal",
                    "Overweight"

@@

            return(cut(
                bmi,
                c(-Inf, 18.5, 22, 25, Inf),

@@

annual$MutClass <- factor(
    annual$MutClass,

@@

annual$NumPulmExacerbation <- cut(
    annual$NumPulmExacerbation,

@@

    breaks = c(0, 1, 2, 3, Inf),

@@

annual$age_group <- cut(
    annual$Age_YrEnd,

@@

    "A_pseudomonasaeruginosa",
    "A_fungalyeast1",
    "A_pulmonarycomplications1",
    "A_mycocultureresults3",
    "A_cultureresults3"

@@

    "A_corticosteroids1",
    "A_feeding2",
    "A_Vx770",
    "A_VX809comb",
    bugs,
    "tube_feeding",
    "liver_disease"

@@

annual[, yn_vars] <- lapply(
    annual[, yn_vars],
    factor,

@@

annual$A_IsOnEnzymes <- factor(
    annual$A_IsOnEnzymes,
    levels = c(1, 0),
    labels = c(" Yes", " No")

@@

annual$ppFEV1_cat <- cut(
    annual$A_FEV1_pct_predicted,
    breaks = c(0, 40, 80, 100, Inf),

@@

annual$ppFEV1_cat <- factor(
    annual$ppFEV1_cat,

@@

    "age_group",
    "Gender",
    "bmi_perc",
    "bmi_cat",
    "MutClass",
    "Race",
    "Hispanicrace",
    "A_IsOnEnzymes",
    "ppFEV1_cat",
    "tube_feeding",
    "NumPulmExacerbation",
    "A_pseudomonasaeruginosa",
    "A_fungalyeast1",
    "A_mycocultureresults3",
    "A_corticosteroids1",
    "cfrd_status_annual",
    "A_Vx770",
    "A_VX809comb",
    "liver_disease"

@@

    cfrd_status_annual = "CFRD Status",
    MutClass = "Mutation class",
    Race = "Race",
    Hispanicrace = "Hispanic",
    Gender = "Sex",
    A_IsOnEnzymes = "Pancreatic enzyme use",

@@

    tube_feeding = "Enteric feedings",
    liver_disease = "Liver disease",

@@

    NumPulmExacerbation = "Admissions for PEx",
    age_group = "Age group",
    height_perc = "Height percentile",
    bmi_cat = "BMI category",

@@

    A_mycocultureresults3 = "Microorganisms in myco culture",
    eDWID = "ID",
    tstart = "Interval start",
    tstop = "Interval stop",
    Age_YrEnd = "Age",
    bmi_perc = "BMI %ile"

@@

# Diagnosis comparisons for HbA1c registry paper

```{r}
# Per reviewer, out of X individuals with a diagnosis of CFRD, X number had an
# OGTT (and X had an A1c, but no OGTT) that year or the year prior.
# This is super slow but idc idc
cfrd_only <- apply(first_cfrd, 1, function(r) {
    id <- as.numeric(r["eDWID"])
    year <- as.numeric(r["first_cfrd"])
    d <- annual[annual$eDWID == id & annual$ReviewYear %in% c(year, year - 1), ]
    a1c <- as.numeric(last(d["A_hgba1c"], na_rm = TRUE))
    ogtt0hr <- as.numeric(last(d["A_ogttfast_bloodglucose"], na_rm = TRUE))
    ogtt2hr <- as.numeric(last(d["A_twohour_bloodglucose"], na_rm = TRUE))
    return(c(id, year, a1c, ogtt0hr, ogtt2hr))
})
cfrd_only <- data.frame(t(cfrd_only))
colnames(cfrd_only) <- c("id", "year", "a1c", "ogtt0hr", "ogtt2hr")
cfrd_only <- left_join(
    cfrd_only,
    annual %>% mutate(eDWID = as.numeric(eDWID)),
    by = join_by(id == eDWID, year == ReviewYear)
)

cfrd_only$a1c_ogtt <- apply(cfrd_only, 1, function(r) {
    not_missing <- paste(
        which(!is.na(r[c("a1c", "ogtt0hr", "ogtt2hr")])),
        collapse = "_"
    )
    return(not_missing)
})
cfrd_only$a1c_ogtt <- factor(
    cfrd_only$a1c_ogtt,
    levels = c("", "1", "1_2", "1_2_3", "1_3", "2", "2_3", "3"),
    labels = c(
        "Neither",
        "HbA1c Only",
        "OGTT",
        "OGTT",
        "OGTT",
        "OGTT",
        "OGTT",
        "OGTT"
    )
)
cfrd_only %>%
    select(
        a1c_ogtt,
        Age_YrEnd,
        A_weight,
        A_height,
        A_bmivalue,
        A_bmipercentile,
        A_FEV1_pct_predicted,
        A_FVC_pct_predicted,
        Race,
        Hispanicrace,
        MutClass,
        a1c,
        A_ogttfast_bloodglucose,
        A_twohour_bloodglucose,
        A_IsOnEnzymes,
        liver_disease,
        A_cultureresults3
    ) %>%
    tbl_summary(
        by = a1c_ogtt,
        missing_text = "Missing",
        statistic = list(all_continuous() ~ "{mean} ({sd})")
    ) %>%
    add_overall() %>%
    add_p(test = list(all_continuous() ~ "oneway.test")) %>%
    modify_caption("**All Groups**")
cfrd_only %>%
    filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) %>%
    mutate(a1c_ogtt = droplevels(a1c_ogtt)) %>%
    select(
        a1c_ogtt,
        Age_YrEnd,
        A_weight,
        A_height,
        A_bmivalue,
        A_bmipercentile,
        A_FEV1_pct_predicted,
        A_FVC_pct_predicted,
        Race,
        Hispanicrace,
        MutClass,
        a1c,
        A_ogttfast_bloodglucose,
        A_twohour_bloodglucose,
        A_IsOnEnzymes,
        liver_disease,
        A_cultureresults3
    ) %>%
    tbl_summary(
        by = a1c_ogtt,
        missing_text = "Missing",
        statistic = list(all_continuous() ~ "{mean} ({sd})")
    ) %>%
    add_overall() %>%
    add_p(test = list(all_continuous() ~ "oneway.test")) %>%
    modify_caption("**OGTT vs. HbA1c Only**")
```

@@

```{r}
data.frame(data_dict$code_meaning[
    data_dict$var %in%
        sub("A_", "", liver_fields)
]) %>%
    gt() %>%
    tab_options(column_labels.hidden = TRUE)
```

@@

    new_cases_10_18 <- sum(inc$n[
        inc$CFRD == "CFRD+" &
            inc$age_group == "[10,19)"
    ])
    at_risk_10_18 <- sum(inc$n[inc$age_group == "[10,19)"])
    inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)

    new_cases_19_30 <- sum(inc$n[
        inc$CFRD == "CFRD+" &
            inc$age_group == "[19,31)"
    ])
    at_risk_19_30 <- sum(inc$n[inc$age_group == "[19,31)"])
    inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)

    new_cases_30 <- sum(inc$n[
        inc$CFRD == "CFRD+" &
            inc$age_group == "[31, Inf)"
    ])
    at_risk_30 <- sum(inc$n[inc$age_group == "[31, Inf)"])
    inc_30 <- new_cases_30 / (at_risk_30 / py)

@@

```{r}
#| message: false
annual %>%
    filter(ReviewYear %in% c(2004, 2011, 2018)) %>%
    select(
        ReviewYear,
        CFRD,
        Age_YrEnd,
        Gender,
        Race,
        Hispanicrace,
        bmi_perc,
        bmi_cat,
        MutClass,
        tube_feeding,
        liver_disease,
        A_FEV1_pct_predicted,
        A_FVC_pct_predicted,
        NumPulmExacerbation,
        A_corticosteroids1,
        A_IsOnEnzymes,
        A_Vx770,
        A_VX809comb
    ) %>%
    tbl_strata(
        strata = ReviewYear,
        .tbl_fun = ~ .x %>%
            tbl_summary(
                by = CFRD,
                missing_text = "Missing",
                type = list(A_corticosteroids1 ~ "categorical"),
                statistic = list(
                    c(
                        "Age_YrEnd",
                        "bmi_perc",
                        "A_FEV1_pct_predicted",
                        "A_FVC_pct_predicted"
                    ) ~
                        "{mean} ({sd})"
                )
            ) %>%
            add_p(
                test = list(
                    c(
                        "Age_YrEnd",
                        "bmi_perc",
                        "A_FEV1_pct_predicted",
                        "A_FVC_pct_predicted"
                    ) ~
                        "t.test"
                )
            ) %>%
            add_overall() %>%
            separate_p_footnotes()
    )
```

@@

```{r}
screening <- annual %>%
    group_by(ReviewYear, age_group) %>%
    summarise(
        `Screened HbA1c` = mean(!is.na(A_hgba1c)),
        `Screened OGTT` = mean(
            !is.na(A_twohour_bloodglucose) |
                !is.na(A_ogttfast_bloodglucose)
        ),
        `Screened Either` = mean(
            !is.na(A_twohour_bloodglucose) |
                !is.na(A_ogttfast_bloodglucose) |
                !is.na(A_hgba1c)
        ),
        .groups = "drop"
    )
screening_table <- screening %>%
    select(ReviewYear, age_group, `Screened OGTT`) %>%
    mutate(`Screened OGTT` = round(`Screened OGTT` * 100, 2)) %>%
    pivot_wider(names_from = age_group, values_from = `Screened OGTT`)
kable(screening_table, caption = "% Screened by OGTT")
screening <- screening %>%
    pivot_longer(`Screened HbA1c`:`Screened Either`)
screening_plot <- ggplot(
    screening,
    aes(
        x = ReviewYear,
        y = value,
        color = age_group
    )
) +
    geom_line() +
    xlab("Year") +
    facet_wrap(~name, nrow = 3) +
    theme_bw() +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_color_discrete(name = "Age Group") +
    theme(text = element_text(size = 20), axis.title.y = element_blank())
screening_plot
ggsave(
    "./Dissemination/Figures/screening_rates.png",
    plot = screening_plot,
    width = 12,
    height = 9,
    units = "in"
)
```

@@

```{r results='asis'}
# How many OGTTs are non-CFRD people getting 2016-2018
ogtts <- annual %>%
    filter(ReviewYear %in% 2016:2018, CFRD == "CFRD-") %>%
    group_by(eDWID) %>%
    summarise(
        `Age Group` = last(na.omit(age_group)),
        `OGTTs` = sum(
            !is.na(A_twohour_bloodglucose) |
                !is.na(A_ogttfast_bloodglucose)
        )
    )
# Add demographics
ogtts$eDWID <- as.numeric(ogtts$eDWID)
ogtts <- left_join(ogtts, demogs, by = join_by(eDWID))
label(ogtts$eDWID) <- "ID"
# Format OGTT counts
ogtts$OGTTs <- factor(
    ogtts$OGTTs,
    levels = 0:3,
    labels = c(
        "No OGTT",
        "At least 1 OGTT",
        "At least 1 OGTT",
        "At least 1 OGTT"
    )
)
# Make table for each variable
age <- ogtts %>%
    select(`Age Group`, OGTTs) %>%
    tbl_summary(by = `Age Group`) %>%
    # add_ci() %>%
    add_p() %>%
    print()
gender <- ogtts %>%
    select(Gender, OGTTs) %>%
    tbl_summary(by = Gender) %>%
    # add_ci() %>%
    add_p() %>%
    print()
muts <- ogtts %>%
    select(MutClass, OGTTs) %>%
    tbl_summary(by = MutClass) %>%
    # add_ci() %>%
    add_p() %>%
    print()
race <- ogtts %>%
    select(Race, OGTTs) %>%
    tbl_summary(by = Race) %>%
    # add_ci() %>%
    add_p() %>%
    print()
eth <- ogtts %>%
    select(Hispanicrace, OGTTs) %>%
    tbl_summary(by = Hispanicrace) %>%
    # add_ci() %>%
    add_p() %>%
    print()
```

@@

```{r}
p <- prev_year_age %>%
    pivot_longer(`[10,19)`:`[31, Inf)`, names_to = "Age Group") %>%
    rename(Year = ReviewYear)
p$name <- "Prevalence of CFRD"
i <- incidence %>%
    select(
        Year,
        `Total Incidence Age 10-18`,
        `Total Incidence Age 19-30`,
        `Total Incidence Age >= 31`
    ) %>%
    pivot_longer(
        `Total Incidence Age 10-18`:`Total Incidence Age >= 31`,
        names_to = "Age Group"
    )
i$name <- "Incidence of CFRD"
s <- screening %>%
    filter(name == "Screened OGTT") %>%
    select(ReviewYear, age_group, value) %>%
    rename(Year = ReviewYear, `Age Group` = age_group)
s$name <- "Screening Rates for CFRD"
s$value <- s$value * 100
t <- rbind(p, i, s)
t$`Age Group` <- factor(
    t$`Age Group`,
    levels = c(
        "[10,19)",
        "[19,31)",
        "[31, Inf)",
        "Total Incidence Age >= 31",
        "Total Incidence Age 10-18",
        "Total Incidence Age 19-30"
    ),
    labels = c(
        "10-18",
        "19-30",
        ">= 31",
        ">= 31",
        "10-18",
        "19-30"
    )
)
t <- t[t$Year >= 2004, ]
# Figure 1
fig1a <- ggplot(
    t[t$name == "Screening Rates for CFRD", ],
    aes(x = Year, y = value, linetype = `Age Group`)
) +
    geom_line() +
    theme_bw() +
    ylab("Percentage") +
    ggtitle("Screening Rates for CFRD")
fig1b <- ggplot(
    t[t$name == "Incidence of CFRD", ],
    aes(x = Year, y = value, linetype = `Age Group`)
) +
    geom_line() +
    theme_bw() +
    ylab("Per 1,000 Person-Years") +
    ggtitle("Incidence of CFRD")
fig1c <- ggplot(
    t[t$name == "Prevalence of CFRD", ],
    aes(x = Year, y = value, linetype = `Age Group`)
) +
    geom_line() +
    theme_bw() +
    ylab("Per 1,000 Person-Years") +
    ggtitle("Prevalence of CFRD")
fig1 <- fig1a +
    fig1b +
    fig1c +
    plot_layout(guides = "collect") +
    plot_annotation(
        title = "Figure 1: Trends in Screening, Incidence, and Prevalence of CFRD from 2003-2018, by age group "
    )
fig1
ggsave(
    "./Dissemination/Figures/fig1.png",
    plot = fig1,
    width = 12,
    height = 3,
    units = "in",
    dpi = 900
)
# Graphical abstract is the same but with color
fig1a <- ggplot(
    t[t$name == "Screening Rates for CFRD", ],
    aes(x = Year, y = value, color = `Age Group`)
) +
    geom_line() +
    theme_bw() +
    ylab("Percentage") +
    ggtitle("Screening Rates for CFRD")
fig1b <- ggplot(
    t[t$name == "Incidence of CFRD", ],
    aes(x = Year, y = value, color = `Age Group`)
) +
    geom_line() +
    theme_bw() +
    ylab("Per 1,000 Person-Years") +
    ggtitle("Incidence of CFRD")
fig1c <- ggplot(
    t[t$name == "Prevalence of CFRD", ],
    aes(x = Year, y = value, color = `Age Group`)
) +
    geom_line() +
    theme_bw() +
    ylab("Per 1,000 Person-Years") +
    ggtitle("Prevalence of CFRD")
fig1 <- fig1a +
    fig1b +
    fig1c +
    plot_layout(guides = "collect") +
    plot_annotation(
        title = "US Cystic Fibrosis Foundation Patient Registry Trends in Screening, Incidence, and Prevalence of Cystic Fibrosis Related Diabetes from 2003-2018"
    )
fig1
ggsave(
    "./Dissemination/Figures/graphical_abstract.png",
    plot = fig1,
    width = 13,
    height = 3,
    units = "in",
    dpi = 900
)
```

@@

```{r results='asis'}
f_vx770 <- update(
    f,
    . ~ . + A_Vx770 - A_corticosteroids1 - NumPulmExacerbation - age_group
)
vx770_cox_res <- coxph(
    f_vx770,
    data = annual[annual$ReviewYear >= 2011, ],
    cluster = eDWID
)
t1 <- vx770_cox_res %>%
    tbl_regression(exponentiate = T)
t1
ggcoef_table(
    vx770_cox_res,
    show_p_values = F,
    add_reference_rows = T,
    exponentiate = T,
    signif_stars = F,
    colour = F
)
```

@@

```{r}
f_vx809comb <- update(
    f,
    . ~
        . +
            A_VX809comb -
            A_corticosteroids1 -
            NumPulmExacerbation -
            age_group -
            Hispanicrace
)
vx809comb_cox_res <- coxph(
    f_vx809comb,
    data = annual[annual$ReviewYear >= 2015, ],
    cluster = eDWID
)
t2 <- vx809comb_cox_res %>%
    tbl_regression(exponentiate = T)
t2
ggcoef_table(
    vx809comb_cox_res,
    show_p_values = F,
    add_reference_rows = T,
    exponentiate = T,
    signif_stars = F,
    colour = F
)
```

@@

```{r}
t1 <- era1 %>% tbl_regression(exponentiate = T)
t1
ggcoef_model(
    era1,
    show_p_values = F,
    add_reference_rows = T,
    exponentiate = T,
    shape_guide = F,
    signif_stars = F,
    colour = "significance",
    shape = "reference_row",
    shape_values = c(16, 15)
)
```
